---
title: "Yield Stability"
author: "Sandesh Shrestha"
format: html
editor: visual
---
#setup
##Loading Packages
```{r Setup}
#| message: false
#| warning: false

library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)
library(leafsync)
library(patchwork)

library(gstat)
library(raster)
library(stars)
library(ggthemes)
library(tibble)
```

```{r}
mytheme_map <- theme_map() +
  theme(legend.position = "right",
        panel.background = element_rect(fill = "lightgrey",
                                        color = NA),
        plot.background = element_rect(fill = "lightgrey",
                                       color = NA),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)
        )
```

##Importing Data
```{r}
yield16 <- read_csv("../Data/Deer Run1 2016 Yield.csv")

yield16
```
#EDA
```{r Summary}
summary(yield16)
```

```{r Yield distribution - Density plot}
ggplot(yield16,
       aes(x = DryYield)) +
  geom_density(fill= "skyblue", alpha =0.5) +
  geom_rug() +
  labs(title = "Raw Yield Density Distribution",
       x = "Yield (bu/ac)",
       y = "Density") +
  theme(plot.title = element_text(hjust = 0.5)) 
  
```
```{r Speed distribution - Density plot}
ggplot(yield16,
       aes(x= speed_mph)) +
  geom_density(fill = "lightpink", alpha = 0.5)+
  geom_rug()
```
```{r Yield map}
ggplot(yield16,
       aes(x = Long,
           y = Lat)) +
  geom_point(aes(color = DryYield)) +
  scale_color_viridis_b() +
mytheme_map
```

```{r Yield  x Speed}
ggplot(yield16,
       aes(x = speed_mph,
           y = DryYield)) +
  geom_point() +
  geom_smooth()
```
#Data Wrangling
```{r}
yield16_w <- yield16 %>%
  clean_names() %>%
  dplyr::select(long, lat,
                speed_mph,
                yield_buac = dry_yield) %>%
  mutate(yield_kgha = yield_buac * 0.453592/0.404686,
         speed_kmh = speed_mph * 1.060934) %>%
  #filter(yield_buac != 0) %>%
  st_as_sf(coords = c("long", "lat")) %>%
  st_set_crs(value = 4326) %>%
  st_transform(crs = 6345)

yield16_w

```
#Maps
```{r}
boundary <- read_sf("../Data/boundary/DR1 boundary.shp")%>%
  st_transform(value = 6345)

boundary
```

```{r }
yield16_w %>%
  filter(yield_buac <400) %>%
  ggplot() +
  geom_sf(aes(color= yield_buac),
          size = .4) +
  scale_color_viridis_b() +
  geom_sf(data = boundary,
          fill = NA,
          size = 2) +
  geom_sf(data = buffer,
          fill = NA,
          color = "red")+
  mytheme_map
```
#EDA
```{r}
summary(yield16_w$yield_buac)
```
## Raw yield quantiles
```{r}
lowq_yield <- quantile(yield16_w$yield_buac, 0.1)
lowq_yield

upq_yield <- quantile(yield16_w$yield_buac, 0.99)
upq_yield

iqr <- upq_yield - lowq_yield
iqr

miny <- lowq_yield - (iqr * 0.25)
miny

maxy <- upq_yield + (iqr * 0.25)
maxy

```

## Min/Max raw yield quantile plot
```{r}
ggplot() +
  geom_density(data = yield16_w,
               aes(x = yield_buac)) +
  geom_vline(xintercept = upq_yield,
             color = "blue",
             linetype = 2
             )+
  geom_vline(xintercept = lowq_yield,
             color = "red",
             linetype = 2
             )+
  geom_vline(xintercept = miny,
             color = "red",
             linetype = 1
             ) +
  geom_vline(xintercept = maxy,
             color = "blue",
             linetype = 1
             )+
  annotate(geom = "rect",
           ymin =0,
           ymax = Inf,
           xmin = miny,
           xmax = maxy,
           alpha = .5,
           fill = "skyblue"
  )
  
```
##Filtering min/max raw yield
```{r}
yield16_w_yf <- yield16_w %>%
  filter(yield_buac >miny &
           yield_buac <maxy)

yield16_w_yf %>% 
  arrange(yield_buac)
```
## Min/Max yield cleaned Summary
```{r}
summary(yield16_w_yf$yield_buac)
summary(yield16_w$yield_buac)
```

# Yield Editor - Min/Max Speed
## Raw Speed Summary
```{r}
summary(yield16_w$speed_mph)
```
## Raw Speed quantiles
```{r}
lowq_speed <- quantile(yield16_w$speed_mph, 0.08)
lowq_speed

upq_speed <- quantile(yield16_w$speed_mph, 0.99)
upq_speed

iqr <- upq_speed - lowq_speed
iqr

minv <- lowq_speed - (iqr * 0.18)
minv

maxv <- upq_speed + (iqr * 0.18)
maxv

```

## Min/Max raw speed quantile plot
```{r}
ggplot() +
  geom_density(data = yield16_w,
               aes(x = speed_mph)) +
  geom_vline(xintercept = upq_speed,
             color = "blue",
             linetype = 2
             )+
  geom_vline(xintercept = lowq_speed,
             color = "red",
             linetype = 2
             )+
  geom_vline(xintercept = minv,
             color = "red",
             linetype = 1
             ) +
  geom_vline(xintercept = maxv,
             color = "blue",
             linetype = 1
             )+
  annotate(geom = "rect",
           ymin =0,
           ymax = Inf,
           xmin = minv,
           xmax = maxv,
           alpha = .5,
           fill = "lightpink"
  )
  
```
##Filtering min/max raw speed
```{r}
yield16_w_sf <- yield16_w_yf %>%
  filter(speed_mph >minv &
           speed_mph <maxv)

yield16_w_sf
```
##Min/Max speed cleaned Summary
```{r}
summary(yield16_w_sf$yield_buac)
summary(yield16_w$speed_mph)
```
```{r}
ggplot(yield16_w_sf,
       aes(x = yield_buac)) +
  geom_density(fill= "skyblue", alpha =0.5) +
  geom_rug() +
  labs(title = "Clean Yield Density Distribution",
       x = "Yield (bu/ac)",
       y = "Density") +
  theme(plot.title = element_text(hjust = 0.5)) 
```


##Field Boundary Plot
### Buffer Boundary
```{r}
buffer <- boundary %>% 
  st_buffer(dist = -40)

```


```{r}
yield16_w_pf <- yield16_w_sf %>% 
  st_intersection(buffer)

yield16_w_pf %>% 
  arrange(yield_buac)
```
##Position Yield Cleaned Summary
```{r}
summary(yield16_w_pf$yield_buac)
```

# Raw vs Cleaned Data
## Sumamry Statistics
```{r}
#Raw
summary_raw <- yield16_w %>%
  summarise( n = length(yield_buac),
             min = min(yield_buac),
             mean = mean(yield_buac),
             max = max(yield_buac),
             sd = sd(yield_buac),
             cv= (sd/mean) * 100
             ) %>%
  mutate(data = "raw") %>%
  dplyr::select(data, everything())

#Filtered
summary_filtered <- yield16_w_sf %>%
  summarise(n = length(yield_buac),
            min = min(yield_buac),
            mean = mean(yield_buac),
            max = max(yield_buac),
            sd = sd(yield_buac),
            cv = (sd/mean)*100
            ) %>%
  mutate(data = "clean") %>%
  dplyr::select(data, everything())

#Combined
summary_raw %>% 
  bind_rows(summary_filtered)

```

## Density Plot
```{r}
ggplot() +
  geom_density(data = yield16_w,
               aes(x= yield_buac),
               fill = "blue",
               alpha = 1
               ) +
  geom_density(data = yield16_w_pf,
               aes(x = yield_buac),
               fill = "yellow",
               alpha = .3)
```
## Raw Yield Map
```{r}
yieldmap_raw <- ggplot(yield16_w) +
  geom_sf(aes(color = yield_buac),
          size = .1) +
   labs(title = "Raw Yield Map",
       vjust = 0.5) +
  theme(axis.text = element_blank(),  # Remove axis text
        axis.ticks = element_blank(), # Remove axis ticks
        axis.title = element_text(hjust = 3), # Remove axis title
        panel.grid = element_blank(), # Remove panel grid
        panel.border = element_blank()) + # Remove panel border
  scale_color_viridis_b()
 

yieldmap_raw
```

## Cleaned Yield Map
```{r}
yieldmap_cleaned <- ggplot() +
  geom_sf(data = yield16_w_pf,
          aes(color = yield_buac),
          size = .2) +
  labs(title = "Cleaned Yield Map") +
  scale_color_viridis_b() +
  theme(axis.text = element_blank(),  # Remove axis text
        axis.ticks = element_blank(), # Remove axis ticks
        axis.title = element_blank(), # Remove axis title
        panel.grid = element_blank(), # Remove panel grid
        panel.border = element_blank()) # Remove panel border

yieldmap_cleaned
```

```{r}
yieldmap_clean_int <- yield16_w_pf %>%
  filter(yield_buac !=0) %>% 
  mapview(zcol = "yield_buac",
        col.regions = viridis(3),
        alpha.regions = 1,
        cex = .1,
        lwd = 0,
        map.types = "Esri.WorldImagery"
)

yieldmap_clean_int
```

## Raw vs Cleaned yield map
```{r}
yieldmap_raw + yieldmap_cleaned
```
#Exporting cleaned version
```{r}
write_sf(yield16_w_pf,
         "../Output/yield_16_clean.geojson",
         delete_dsn = T)
```

# Yield Interpolation
```{r}
yield_c <- yield16_w_pf
yield_c
```

## Grid
```{r}
grid_v <-  st_make_grid(boundary,
                        cellsize = 10
                        )
grid_v
```
```{r grid vector plot}
yieldmap_cleaned +
  geom_sf(data = grid_v,
          alpha = 0.5,
          fill = NA,
          size = .1)

```
## Rasterize our vector grid
```{r grid_r}
grid_r <- grid_v %>%
  st_as_sf() %>% 
  st_rasterize(dx = 10,
               dy = 10) %>% 
  st_crop(boundary)

grid_r
```
```{r}
yieldmap_cleaned +
  geom_stars(data = grid_r,
             alpha = .6)
```
## Inverse Distance Weight (IDW) 
```{r}
idw_mod <- gstat(formula = yield_buac ~1,
                 data = yield_c
                 )
idw_mod
```
```{r idw pred}
idw_pred <- predict(idw_mod,
                    grid_r)
idw_pred
```
```{r idw_pred plot}
ggplot() +
  geom_stars(data = idw_pred) +
  scale_fill_viridis() +
  geom_sf(data = boundary,
          fill = NA) +
  geom_sf(data = grid_v,
          alpha = .5,
          fill = NA,
          size = .1)
```
```{r idw_pred_v}
idw_pred_v <- idw_pred %>% 
  st_as_sf(as_points = F,
           merge = F) %>% 
  dplyr::select(ipyield_buac = var1.pred)

idw_pred_v
```

```{r idw_pred_v plot}
ggplot()+
  geom_sf(data = idw_pred_v,
          aes(fill = ipyield_buac),
          color = NA) +
  scale_fill_viridis_b(name = "Yield\n(bu/ac)",
                       breaks = seq(0, 250, by = 50)) +
  labs(title = "Cleaned Interpolated Yield Map")+
  geom_sf(data = boundary,
          fill = NA) +
mytheme_map
```

```{r interpolated distribution}
ggplot()+
  geom_density(data = idw_pred_v,
               aes(x = ipyield_buac)
               ) +
  geom_density(data = yield_c,
               aes(x = yield_buac),
               color = "blue")

```
# Exporting Interpolated Data
```{r Interpolated Yield Data}
write_sf(idw_pred_v,
         "../Output/yield_interpolated.geojson",
         delete_dsn = T)
```







